Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-06-25 paper/paper.Rmd
add spatiotemporal intro

## Spatial or Spatiotemporal

<!-- FIXME: explain that most methdso can also be spatial-only -->

@@

Group-level CV follows the principle of an inherit grouping structure of the observations, both in space and/or time, to determine (groups of) partitions.
In contrast to blocking or clustering not the the spatial or temporal location is used but additional metadata such as the membership to a specific crop type [@brenning2006].
The idea to follow a such an inherited grouping structure of the data is also referred to as "target-oriented validation" [@cast].
Even though flexibility is limited due to the inherited grouping structure within the data, some implementations allow to specify the resulting number of folds by omitting parts of the dataset.

\pkg{mlr3spatiotempcv} supports two "grouping" implementations: (1) The "Leave-location-and-time-out" family from package \pkg{CAST} and (2) "Custom Resampling" from \pkg{mlr3}.

@@

In this example the `cookfarm` dataset is used because it contains a temporal variable ("Date") which is not present in dataset `ecuador`.

Two visualization types are supported: a two-dimensional and a three-dimensional method which can be selected via the logical argument `plot3d`.
The 3D visualization (space + time) has been implemented via \pkg{plotly}.
Because a dynamic image cannot be included in this manuscript, a static version, which can be generated by setting `static_image = TRUE`, is shown.

In \pkg{mlr3spatiotempcv} the method is named `"sptcv_cstf"` because the underlying upstream implementation in package \pkg{CAST} is named `CreateSpacetimeFolds()`.

@@

The following code uses the `cookfarm` dataset as the base, modifies to the `Date` variable to contain five unique levels and then creates a spatiotemporal regression task in \pkg{mlr3spatiotempcv} (Fig.\@ref(fig:lto)).

@@

Method "Leave-Location-Out" (LLO) works similar to "Leave-Time-Out" (LTO) in that it operates in a two-dimensional space.
Here, specific locations are left out for forming the partitions while the temporal information is ignored (Fig.\@ref(fig:llo)).

@@

The "Leave-Location-and-Time-Out" (LLTO) method combines both "Leave-Time-Out" (LTO) and "Leave-Location-Out" (LLO).
It creates partitions that operate on spatiotemporal clusters to reduce both spatial and temporal autocorrelation.
To achieve this, observations are omitted and only a subset is used.
In the special scenario with five temporal levels and five folds, the train/test split happens only at one temporal level (Fig.\@ref(fig:llto)).

>>>>

## Spatial, temporal or spatiotemporal

Some methods have the ability to operate in multiple spaces, i.e. spatial, temporal or spatiotemporal (2D + time).
To avoid redundancy and putting too much weight on certain methods, only the spatiotemporal examples of these methods will be shown in the following sections.
However, these can be adapted for the respective other cases (spatial only, temporal only).

@@

Group-level CV follows the principle of an inherit grouping structure of the observations, both in space and/or time, to determine (groups of) partitions.
In contrast to blocking or clustering not the the spatial or temporal location is used but additional metadata such as the membership to a specific crop type [@brenning2006].
The idea to follow an inherited grouping structure of the data is also referred to as "target-oriented validation" [@cast].
Even though flexibility with respect to fold creation is limited due to the inherited grouping structure within the data, some implementations allow to specify the resulting number of folds by omitting parts of the dataset.

\pkg{mlr3spatiotempcv} supports two "grouping" implementations: (1) The "Leave-location-and-time-out" family from package \pkg{CAST} and (2) "Custom Resampling" from the \pkg{mlr3} package.

@@

In this example the `cookfarm` dataset is used because it contains the variable ("Date") which serves as the temporal component in the dataset.

Two visualization types are supported by `mlr3spatiotempcv::autoplot()` for this method: a two-dimensional and a three-dimensional one which can be selected via the logical argument `plot3D`.
The 3D visualization (i.e. 2D + time) is done via package \pkg{plotly}.
Because a dynamic image cannot be included in this manuscript, a static version, which can be generated by setting `static_image = TRUE`, is shown.

In \pkg{mlr3spatiotempcv} the method is named `"sptcv_cstf"` which is an acronym originating from the underlying upstream implementation in package \pkg{CAST} `CreateSpacetimeFolds()`.

@@

The following code uses the `cookfarm` dataset as the base, modifies to the `Date` variable to contain five unique levels and then creates a spatiotemporal regression task in \pkg{mlr3spatiotempcv} (Fig. \@ref(fig:lto)).

@@

Method "Leave-Location-Out" (LLO) works similar to "Leave-Time-Out" (LTO) in that it operates in a two-dimensional space.
Here, specific locations are left out for forming the partitions while the temporal information is ignored (Fig. \@ref(fig:llo)).

@@

The "Leave-Location-and-Time-Out" (LLTO) method combines both "Leave-Time-Out" (LTO) and "Leave-Location-Out" (LLO).
It creates partitions that operate on spatiotemporal clusters to reduce both spatial and temporal autocorrelation.
To achieve this, observations are omitted and only a subset is used.
In the special scenario with five temporal levels and five folds, the train/test split happens only at one temporal level (Fig. \@ref(fig:llto)).
